### 2015-01-26 p3/submission.Rmd
add univariate and bivariate plots and analysis

Exploring Red Wine Quality by Allan Reyes
========================================================
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(GGally)
library(scales)
library(memisc)
library(reshape)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
df <- read.csv('data/wineQualityReds.csv')
```

## About

In this exercise, I will explore a data set on wine quality and physicochemical properties.  The objective is to explore which chemical properties influence the quality of red wines.  I'll start by exploring the data using the statistical program, R.  As interesting relationships in the data are discovered, I'll produce and refine plots to illustrate them.  The data is available for download [here](https://www.google.com/url?q=https%3A%2F%2Fs3.amazonaws.com%2Fudacity-hosted-downloads%2Fud651%2FwineQualityReds.csv&sa=D&sntz=1&usg=AFQjCNFbtvciXJWwSbQuwE2Br_pH9JquAg) and background information is available at this [link](https://www.google.com/url?q=https%3A%2F%2Fs3.amazonaws.com%2Fudacity-hosted-downloads%2Fud651%2FwineQualityInfo.txt&sa=D&sntz=1&usg=AFQjCNHZWanxQ_JGIKpDr2lo9rcDF9jBWQ).

## Summary Statistics
Let's run some basic functions to examine the structure and schema of the data set.

```{r echo=TRUE}
str(df)
summary(df)
```

Since we're primarily interested in quality, it would also be interesting to see basic statistics on that as well.

```{r echo=TRUE}
summary(df$quality)
```

Some initial observations here:

- There are 1599 observations of 13 numeric variables.
- `X` appears to be the unique identifier.
- `quality` is an ordered, categorical, discrete variable.  From the literature, this was on a 0-10 scale, and was rated by at least 3 wine experts.  The values ranged only from 3 to 8, with a mean of 5.6 and median of 6.
- All other variables seem to be continuous quantities (w/ the exception of the `.sulfur.dioxide` suffixes).
- From the variable descriptions, it appears that `fixed.acidity` ~ `volatile.acidity` and `free.sulfur.dioxide` ~ `total.sulfur.dioxide` may possible by dependent, subsets of each other.

```{r echo=FALSE, results='hide', Initial_Observations}
# let's at least explore, clean up, and format the first two points.
# X
df$X = factor(df$X)

# quality
summary(df$quality)
table(df$quality)

# assertion was correct here, so let's ensure the data frame semantically
# reflects that.
df$quality <- factor(df$quality, ordered = T)
str(df$quality)
```

## Univariate Plots

To first explore this data visually, I'll draw up quick histograms of all 12 variables.  The intention here is to see a quick distribution of the values.

```{r echo=FALSE, message=FALSE, Univariate_Plots}
# exploratory, quick histogram plots
grid.arrange(qplot(df$fixed.acidity),
             qplot(df$volatile.acidity),
             qplot(df$citric.acid),
             qplot(df$residual.sugar),
             qplot(df$chlorides),
             qplot(df$free.sulfur.dioxide),
             qplot(df$total.sulfur.dioxide),
             qplot(df$density),
             qplot(df$pH),
             qplot(df$sulphates),
             qplot(df$alcohol),
             qplot(df$quality),
             ncol = 4)
```

## Univariate Analysis

### Wine Quality
I first looked at wine quality.  Although it has a discrete range of only 3-8, we can roughly see that there is some amount of normal distribution.  A large majority of the wines examined received ratings of 5 or 6, and very few received 3, 4, or 8.  There's not much more we can do with this histogram, as both decreasing or increasing bin sizes would distort the data.

Given the ratings and distribution of wine quality, I'll instantiate another categorical variable, classifying the wines as 'bad' (rating 0 to 4), 'average' (rating 5 or 6), and 'good' (rating 7 to 10).

```{r echo=FALSE, message=FALSE}
df$rating <- ifelse(df$quality < 5, 'bad', ifelse(df$quality < 7, 'average', 'good'))
df$rating <- ordered(df$rating, levels = c('bad', 'average', 'good'))
summary(df$rating)
```

```{r echo=FALSE, message=FALSE}
qplot(df$rating)
```

### Distributions and Outliers
- It appears that density and pH are normally distributed, with few outliers.
- Fixed and volatile acidity, sulfur dioxides, sulphates, and alcohol seem to be long-tailed.
- Qualitatively, residual sugar and chlorides have extreme outliers.
- Citric acid appeared to have a large number of zero values.  I'm curious whether this is truly zero, or if it is a case of non-reporting.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = df,
       aes(x = fixed.acidity)) +
  geom_histogram() +
  scale_x_log10()

ggplot(data = df,
       aes(x = volatile.acidity)) +
  geom_histogram() +
  scale_x_log10()

ggplot(data = df,
       aes(x = citric.acid)) +
  geom_histogram() +
  scale_x_log10()
```

When plotted on a base 10 logarithmic scale, `fixed.acidity` and `volatile.acidity` appear to be normally-distributed.  This makes sense, considering that `pH` is normally distributed, and pH, by definition, is a measure of acidity and is on a logarithmic scale.  Curiously, however, `citric.acid`, did not appear to be normally-distributed on a logarithmic scale.  Upon further investigation:

```{r}
length(subset(df, citric.acid == 0)$citric.acid)
```

It is apparent that 132 observations had a value of zero.  This yields some concerns on whether or not these 132 values were reported or not, considering that the next 'bin' higher contains only 32 observations.

```{r echo=FALSE, message=FALSE}
p1 <- ggplot(data = df, aes(x = residual.sugar)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0, quantile(df$residual.sugar, 0.95))) +
  xlab('residual.sugar, 95th percentile truncated')

p2 <- p1 + scale_x_log10() + xlab('residual.sugar, log10')
grid.arrange(p1, p2, ncol=1)

p1 <- ggplot(data = df, aes(x = chlorides)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0, quantile(df$chlorides, 0.95))) +
  xlab('chlorides, 95th percentile truncated')

p2 <- p1 + scale_x_log10() + xlab('chlorides, log10')
grid.arrange(p1, p2, ncol=1)

p1 <- ggplot(data = df, aes(x = sulphates)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0, quantile(df$sulphates, 0.95))) +
  xlab('sulphates, 95th percentile truncated')

p2 <- p1 + scale_x_log10() + xlab('sulphates, log10')
grid.arrange(p1, p2, ncol=1)

rm(p1, p2)
```

From these plots, it appears that chlorides and sulphates have a small number of high outliers.  Residual sugar, however, appears to be long-tailed.  I'll instatiate a separate variable classifier, here, to annotate which wines are 'sweet'.  Using a (temporary) threshold of

```{r}
subset(df, residual.sugar > 45)
```

### Short questions

#### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

While exploring the univariate histogram distributions, there did not appear to be any bimodal or multimodal distributions that would warrant sub-classification into categorical variables.  I considered potentially splitting `residual.sugar` into 'sweet wine' and 'dry wine', but the [Wikipedia source](http://en.wikipedia.org/wiki/Sweetness_of_wine#Residual_sugar) cited a residual sugar of greater than 45 g/L or g/m^3 to classify as a sweet wine.

#### Did you create any new variables from existing variables in the dataset?

I instantiated an ordered factor, `rating`, classifying each wine sample as 'bad', 'average', or 'good'.

Upon further examination of the data set documentation, it appears that `fixed.acidity` and `volatile.acidity` are different types of acids; tartaric acid and acetic acid.  I decided to create a combined variable, `TAC.acidity`, containing the sum of tartaric, acetic, and citric acid.

```{r echo=FALSE}
df$TAC.acidity <- df$fixed.acidity + df$volatile.acidity + df$citric.acid
qplot(df$TAC.acidity)
```

#### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

I addressed the distributions in the 'Distributions' section.  Boxplots are better suited in visualizing the outliers.

```{r echo=FALSE}
grid.arrange(qplot(data = df, x = 'boxplot',
                   y = fixed.acidity,geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = volatile.acidity, geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = citric.acid, geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = TAC.acidity, geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = residual.sugar, geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = chlorides, geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = free.sulfur.dioxide, geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = total.sulfur.dioxide, geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = density, geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = pH, geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = sulphates, geom = 'boxplot'),
             qplot(data = df, x = 'boxplot',
                   y = alcohol, geom = 'boxplot'),
             ncol = 4)
```

In univariate analysis, I chose not to tidy or adjust any data, short of plotting a select few on logarithmic scales.  Bivariate boxplots, with X as rating or quality, will be more interesting in showing trends with wine quality.

## Bivariate Plots and Analysis
```{r echo=FALSE, Summary_Statistics}
# ggpairs was WAY too slow!  Uncomment and use at your own risk!
#set.seed(1)
#df_sample <- df[,-which(names(df) %in% c('X', 'rating'))][sample(1:length(df$quality), 40), ]
#ggpairs(df_sample, params = c(shape = I('.'), outlier.shape = I('.')))
```

To get a quick snapshot of how the variables affect quality, I generated box plots for each.

```{r echo=FALSE, Bivariate_Plots}
grid.arrange(qplot(data = df, x = quality,
                   y = fixed.acidity, geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = volatile.acidity, geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = citric.acid, geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = TAC.acidity, geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = log10(residual.sugar), geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = log10(chlorides), geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = free.sulfur.dioxide, geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = total.sulfur.dioxide, geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = density, geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = pH, geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = log10(sulphates), geom = 'boxplot'),
             qplot(data = df, x = quality,
                   y = alcohol, geom = 'boxplot'),
             ncol = 4)

grid.arrange(qplot(data = df, x = rating,
                   y = fixed.acidity, geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = volatile.acidity, geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = citric.acid, geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = TAC.acidity, geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = log10(residual.sugar), geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = log10(chlorides), geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = free.sulfur.dioxide, geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = total.sulfur.dioxide, geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = density, geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = pH, geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = log10(sulphates), geom = 'boxplot'),
             qplot(data = df, x = rating,
                   y = alcohol, geom = 'boxplot'),
             ncol = 4)
```

From exploring these plots, it seems that a 'good' wine generally has these trends:

- higher fixed acidity (tartaric acid) and citric acid, lower volatile acidity (acetic acid)
- lower pH (i.e. more acidic)
- higher sulphates
- higher alcohol
- to a lesser extend, lower chlorides and lower density

Residual sugar and sulfur dioxides did not seem to have a dramatic impact on the quality or rating of the wines.  Interestingly, it appears that different types of acid affect wine quality different; as such, `TAC.acidity` saw an attenuated trend, as the presence of volatile (acetic) acid accompanied decreased quality.

By utilizing `cor.test`, I calculated the correlation for each of these variables against `quality`:

```{r echo=FALSE}
correlations <- c(
cor.test(df$fixed.acidity, as.numeric(df$quality))$estimate,
cor.test(df$volatile.acidity, as.numeric(df$quality))$estimate,
cor.test(df$citric.acid, as.numeric(df$quality))$estimate,
cor.test(df$TAC.acidity, as.numeric(df$quality))$estimate,
cor.test(log10(df$residual.sugar), as.numeric(df$quality))$estimate,
cor.test(log10(df$chlorides), as.numeric(df$quality))$estimate,
cor.test(df$free.sulfur.dioxide, as.numeric(df$quality))$estimate,
cor.test(df$total.sulfur.dioxide, as.numeric(df$quality))$estimate,
cor.test(df$density, as.numeric(df$quality))$estimate,
cor.test(df$pH, as.numeric(df$quality))$estimate,
cor.test(log10(df$sulphates), as.numeric(df$quality))$estimate,
cor.test(df$alcohol, as.numeric(df$quality))$estimate)
names(correlations) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',
                         'TAC.acidity', 'log10.residual.sugar',
                         'log10.chlordies', 'free.sulfur.dioxide',
                         'total.sulfur.dioxide', 'density', 'pH',
                         'log10.sulphates', 'alcohol')
correlations
```

Quantitatively, it appears that the following variables have relatively higher correlations to wine quality:

- alcohol
- sulphates (log10)
- volatile acidity
- citric acid

Let's see how these variables compare, plotted against each other and faceted by wine `rating`:

```{r echo=FALSE}
ggplot(data = df, aes(x = log10(sulphates), y = alcohol)) +
  facet_wrap(~rating) +
  geom_point()

ggplot(data = df, aes(x = volatile.acidity, y = alcohol)) +
  facet_wrap(~rating) +
  geom_point()

ggplot(data = df, aes(x = citric.acid, y = alcohol)) +
  facet_wrap(~rating) +
  geom_point()

ggplot(data = df, aes(x = volatile.acidity, y = log10(sulphates))) +
  facet_wrap(~rating) +
  geom_point()

ggplot(data = df, aes(x = citric.acid, y = log10(sulphates))) +
  facet_wrap(~rating) +
  geom_point()

ggplot(data = df, aes(x = citric.acid, y = volatile.acidity)) +
  facet_wrap(~rating) +
  geom_point()
```

The relative value of these scatterplots are suspect; if anything, it illustrates how heavily alcohol content affects rating.  The weakest bivariate relationship appeared to be alcohol vs. citric acid.  The plots were nearly uniformly-distributed.  The strongest relationship appeared to be volatile acididty vs. citric acid, which had a negative correlation.

Examining the acidity variables, I saw strong correlations between them:

```{r echo=FALSE}
ggplot(data = df, aes(x = fixed.acidity, y = citric.acid)) +
  geom_point()
cor.test(df$fixed.acidity, df$citric.acid)

ggplot(data = df, aes(x = volatile.acidity, y = citric.acid)) +
  geom_point()
cor.test(df$volatile.acidity, df$citric.acid)

ggplot(data = df, aes(x = log10(TAC.acidity), y = pH)) +
  geom_point()
cor.test(log10(df$TAC.acidity), df$pH)
```

Most notably, base 10 logarithm `TAC.acidity` correlated very well with pH.  This is certainly expected, as pH is essentially a measure of acidity.  An interesting question to pose, using basic chemistry knowledge, is to ask what other components other than the measured acids are affecting pH.  We can quantify this difference by building a predictive linear model, to predict pH based off of `TAC.acidity` and capture the % difference as a new variable.

```{r echo=FALSE}
m <- lm(I(pH) ~ I(log10(TAC.acidity)), data = df)
df$pH.predictions <- predict(m, df)
# (observed - expected) / expected
df$pH.error <- (df$pH.predictions - df$pH)/df$pH

ggplot(data = df, aes(x = quality, y = pH.error)) +
  geom_boxplot()
```

The median % error hovered at or near zero for most wine qualities.  Notably, wines rated with a quality of 3 had large negative error.  We can interpret this finding by saying that for many of the 'bad' wines, total acidity from tartaric, acetic, and citric acids were a worse predictor of pH.  Simply put, it is likely that there were other components--possibly impurities--that changed and affected the pH.

As annotated previously, I hypothesized that `free.sulfur.dioxide` and `total.sulfur.dioxide` were dependent on each other.  Plotting this:

```{r echo=FALSE}
ggplot(data = df, aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide)) +
  geom_point() +
  geom_smooth()

cor.test(df$free.sulfur.dioxide, df$total.sulfur.dioxide)
```

It is clear that there is a very strong relationship between the two.  Aside from `TAC.acidity`, this seemed to be the strongest bivariate relationship.  Additionally, despite the telling name descriptions, the clear 'floor' on this graph hints that `free.sulfur.dioxide` is a subset of `total.sulfur.dioxide`.

## Multivariate Plots

```{r echo=FALSE, Multivariate_Plots}

```

## Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

## Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One

### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two

### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

### 2015-01-26 p3/submission.Rmd
add multivariate plots and analysis

@@
 ```{r echo=FALSE, Multivariate_Plots}
+ggplot(data = df,
+       aes(x = citric.acid, y = volatile.acidity,
+           color = quality)) +
+  geom_point() +
+  facet_wrap(~rating)
 
+ggplot(data = df,
+       aes(x = alcohol, y = log10(sulphates),
+           color = quality)) +
+  geom_point() +
+  facet_wrap(~rating)
+
+ggplot(data = df,
+       aes(x = pH, y = alcohol, color = quality)) +
+  geom_point() +
+  facet_wrap(~rating)
 ```
@@
 
-### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
-
-### Were there any interesting or surprising interactions between features?
-
-### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.
-
-------
+I primarily examined the 4 features which showed high correlation with quality.
+These scatterplots were a bit crowded, so I faceted by `rating` to illustrate the population differences between good wines, average wines, and bad wines.
+It's clear that a higher citric acid and lower volatile (acetic) acid contributes towards better wines.
+Likewise, better wines tended to have higher sulphates and alcohol content.
+Surprisingly, pH had very little visual impact on wine quality, and was shadowed by the larger impact of alcohol.
+Interestingly, this shows that what makes a good wine depends on the type of acids that are present.
 

### 2015-01-26 p3/submission.Rmd
add reflection

@@
 
-### Plot One
+### Effect of acids on wine quality
 ```{r echo=FALSE, Plot_One}
+grid.arrange(ggplot(data = df, aes(x = quality, y = fixed.acidity,
+                                   fill = quality)) + 
+               ggtitle('Fixed Acidity in Different Wine Quality') +
+               ylab('Fixed Acidity (g/dm^3)') +
+               xlab('Quality') +
+               geom_boxplot(),
+             ggplot(data = df, aes(x = quality, y = volatile.acidity,
+                                   fill = quality)) +
+               ggtitle('Volatile Acidity in Different Wine Quality') +
+               ylab('Volatile Acidity (g/dm^3)') +
+               xlab('Quality') +
+               geom_boxplot(), 
+             ggplot(data = df, aes(x = quality, y = citric.acid,
+                                   fill = quality)) +
+               ggtitle('Citric Acidity in Different Wine Quality') +
+               ylab('Citric Acid (g/dm^3)') +
+               xlab('Quality') +
+               geom_boxplot(), 
+             ggplot(data = df, aes(x = quality, y = pH,
+                                   fill = quality)) +
+               ggtitle('pH in Different Wine Quality') +
+               ylab('pH') +
+               xlab('Quality') +
+               geom_boxplot())
 
@@
 
-### Description One
+These subplots were created to demonstrate the effect of acidity and pH on wine quality.
+Generally, higher acidity (or lower pH) is seen in highly-rated wines.
+To caveat this, a presence of volatile (acetic) acid negatively affected wine quality.
+Citric acidity had a high correlation with wine quality, while fixed (tartaric) acid had a smaller impact.
 
-### Plot Two
+### Effect of Alcohol on Wine Quality
 ```{r echo=FALSE, Plot_Two}
-
+ggplot(data = df, aes(x = quality, y = alcohol,
+                      fill = rating)) +
+  geom_boxplot() +
+  ggtitle('Alcohol Levels in Different Wine Qualities') +
+  xlab('Quality') +
+  ylab('Alcohol (% volume)')
 ```
 
-### Description Two
+These boxplots demonstrate the effect of alcohol content on wine quality.
+Generally, higher alcohol content correlated with higher wine quality.
+However, as the outliers and intervals show, alchol content alone did not produce a higher quality.
 
@@
 ```{r echo=FALSE, Plot_Three}
-
+ggplot(data = subset(df, rating != 'average'), aes(x = volatile.acidity, y = alcohol,
+                      color = rating)) +
+  geom_point() +
+  ggtitle('Alcohol vs. Volatile Acidity and Wine Quality') +
+  xlab('Volatile Acidity (g / dm^3)') +
+  ylab('Alcohol (% volume)')
 ```
 
-### Description Three
-
-------
+This is perhaps the most telling graph.
+I subsetted the data to remove the 'average' wines, or any wine with a rating of 5 or 6.
+As the correlation tests show, wine quality was affected most strongly by alcohol and volaticle acidity.
+While the boundaries are not as clear cut or modal, it's apparent that high volatile acidity--with few exceptions--kept wine quality down.
+A combination of high alcohol content and low volatile acidity produced better wines.
 
 # Reflection
+Through this exploratory data analysis, I was able to identify the key factors that determine and drive wine quality, mainly: alcohol content, sulphates, and acidity.
+It is important to note, however, that wine quality is ultimately a subjective measure, albeit measured by wine experts.
+That said, the correlations for these variables are within reasonable bounds.
+The graphs adequately illustrate the factors that make good wines 'good' and bad wines 'bad'.
+Further study with inferential statistics could be done to quantitatively confirm these assertions.

### 2015-01-26 p3/submission.Rmd
add functions

@@
 From these plots, it appears that chlorides and sulphates have a small number of high outliers.
-Residual sugar, however, appears to be long-tailed.
-I'll instatiate a separate variable classifier, here, to annotate which wines are 'sweet'.
-Using a (temporary) threshold of
-
-```{r}
-subset(df, residual.sugar > 45)
-```
 
@@
 ```{r echo=FALSE}
-grid.arrange(qplot(data = df, x = 'boxplot',
-                   y = fixed.acidity,geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = volatile.acidity, geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = citric.acid, geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = TAC.acidity, geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = residual.sugar, geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = chlorides, geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = free.sulfur.dioxide, geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = total.sulfur.dioxide, geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = density, geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = pH, geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = sulphates, geom = 'boxplot'),
-             qplot(data = df, x = 'boxplot',
-                   y = alcohol, geom = 'boxplot'),
+get_simple_boxplot <- function(column) {
+  return(qplot(data = df, x = 'simple',
+               y = column, geom = 'boxplot',
+               xlab = ''))
+}
+
+grid.arrange(get_simple_boxplot(df$fixed.acidity),
+             get_simple_boxplot(df$volatile.acidity),
+             get_simple_boxplot(df$citric.acid),
+             get_simple_boxplot(df$TAC.acidity),
+             get_simple_boxplot(df$residual.sugar),
+             get_simple_boxplot(df$chlorides),
+             get_simple_boxplot(df$free.sulfur.dioxide),
+             get_simple_boxplot(df$total.sulfur.dioxide),
+             get_simple_boxplot(df$density),
+             get_simple_boxplot(df$pH),
+             get_simple_boxplot(df$sulphates),
+             get_simple_boxplot(df$alcohol),
              ncol = 4)
@@
 ```{r echo=FALSE, Bivariate_Plots}
-grid.arrange(qplot(data = df, x = quality,
-                   y = fixed.acidity, geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = volatile.acidity, geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = citric.acid, geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = TAC.acidity, geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = log10(residual.sugar), geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = log10(chlorides), geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = free.sulfur.dioxide, geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = total.sulfur.dioxide, geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = density, geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = pH, geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = log10(sulphates), geom = 'boxplot'),
-             qplot(data = df, x = quality,
-                   y = alcohol, geom = 'boxplot'),
+get_bivariate_boxplot <- function(x, y, xlab, ylab) {
+  return(qplot(data = df, x = x, y = y, geom = 'boxplot'))
+}
+
+grid.arrange(get_bivariate_boxplot(df$quality, df$fixed.acidity),
+             get_bivariate_boxplot(df$quality, df$volatile.acidity),
+             get_bivariate_boxplot(df$quality, df$citric.acid),
+             get_bivariate_boxplot(df$quality, df$TAC.acidity),
+             get_bivariate_boxplot(df$quality, log10(df$residual.sugar)),
+             get_bivariate_boxplot(df$quality, log10(df$chlorides)),
+             get_bivariate_boxplot(df$quality, df$free.sulfur.dioxide),
+             get_bivariate_boxplot(df$quality, df$total.sulfur.dioxide),
+             get_bivariate_boxplot(df$quality, df$density),
+             get_bivariate_boxplot(df$quality, df$pH),
+             get_bivariate_boxplot(df$quality, log10(df$sulphates)),
+             get_bivariate_boxplot(df$quality, df$alcohol),
              ncol = 4)
 
-grid.arrange(qplot(data = df, x = rating,
-                   y = fixed.acidity, geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = volatile.acidity, geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = citric.acid, geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = TAC.acidity, geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = log10(residual.sugar), geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = log10(chlorides), geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = free.sulfur.dioxide, geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = total.sulfur.dioxide, geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = density, geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = pH, geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = log10(sulphates), geom = 'boxplot'),
-             qplot(data = df, x = rating,
-                   y = alcohol, geom = 'boxplot'),
+grid.arrange(get_bivariate_boxplot(df$rating, df$fixed.acidity),
+             get_bivariate_boxplot(df$rating, df$volatile.acidity),
+             get_bivariate_boxplot(df$rating, df$citric.acid),
+             get_bivariate_boxplot(df$rating, df$TAC.acidity),
+             get_bivariate_boxplot(df$rating, log10(df$residual.sugar)),
+             get_bivariate_boxplot(df$rating, log10(df$chlorides)),
+             get_bivariate_boxplot(df$rating, df$free.sulfur.dioxide),
+             get_bivariate_boxplot(df$rating, df$total.sulfur.dioxide),
+             get_bivariate_boxplot(df$rating, df$density),
+             get_bivariate_boxplot(df$rating, df$pH),
+             get_bivariate_boxplot(df$rating, log10(df$sulphates)),
+             get_bivariate_boxplot(df$rating, df$alcohol),
              ncol = 4)
@@
 ```{r echo=FALSE}
+simple_cor_test <- function(x, y) {
+  return(cor.test(x, y)$estimate)
+}
+
 correlations <- c(
-cor.test(df$fixed.acidity, as.numeric(df$quality))$estimate,
-cor.test(df$volatile.acidity, as.numeric(df$quality))$estimate,
-cor.test(df$citric.acid, as.numeric(df$quality))$estimate,
-cor.test(df$TAC.acidity, as.numeric(df$quality))$estimate,
-cor.test(log10(df$residual.sugar), as.numeric(df$quality))$estimate,
-cor.test(log10(df$chlorides), as.numeric(df$quality))$estimate,
-cor.test(df$free.sulfur.dioxide, as.numeric(df$quality))$estimate,
-cor.test(df$total.sulfur.dioxide, as.numeric(df$quality))$estimate,
-cor.test(df$density, as.numeric(df$quality))$estimate,
-cor.test(df$pH, as.numeric(df$quality))$estimate,
-cor.test(log10(df$sulphates), as.numeric(df$quality))$estimate,
-cor.test(df$alcohol, as.numeric(df$quality))$estimate)
+  simple_cor_test(df$fixed.acidity, as.numeric(df$quality)),
+  simple_cor_test(df$volatile.acidity, as.numeric(df$quality)),
+  simple_cor_test(df$citric.acid, as.numeric(df$quality)),
+  simple_cor_test(df$TAC.acidity, as.numeric(df$quality)),
+  simple_cor_test(log10(df$residual.sugar), as.numeric(df$quality)),
+  simple_cor_test(log10(df$chlorides), as.numeric(df$quality)),
+  simple_cor_test(df$free.sulfur.dioxide, as.numeric(df$quality)),
+  simple_cor_test(df$total.sulfur.dioxide, as.numeric(df$quality)),
+  simple_cor_test(df$density, as.numeric(df$quality)),
+  simple_cor_test(df$pH, as.numeric(df$quality)),
+  simple_cor_test(log10(df$sulphates), as.numeric(df$quality)),
+  simple_cor_test(df$alcohol, as.numeric(df$quality)))
 names(correlations) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',

### 2015-01-26 p3/submission.Rmd
small update + update html
add breaks to keep lines under 80 characters

@@
 ```{r echo=FALSE, message=FALSE}
-df$rating <- ifelse(df$quality < 5, 'bad', ifelse(df$quality < 7, 'average', 'good'))
-df$rating <- ordered(df$rating, levels = c('bad', 'average', 'good'))
+df$rating <- ifelse(df$quality < 5, 'bad', ifelse(
+  df$quality < 7, 'average', 'good'))
+df$rating <- ordered(df$rating,
+                     levels = c('bad', 'average', 'good'))
 summary(df$rating)
@@
 ```
-
-From these plots, it appears that chlorides and sulphates have a small number of high outliers.
 
@@
 ```{r echo=FALSE, Plot_Three}
-ggplot(data = subset(df, rating != 'average'), aes(x = volatile.acidity, y = alcohol,
+ggplot(data = subset(df, rating != 'average'),
+       aes(x = volatile.acidity, y = alcohol,
                       color = rating)) +

### 2015-01-26 p3/submission.Rmd
add label fixes

@@
 ```{r echo=FALSE}
-get_simple_boxplot <- function(column) {
+get_simple_boxplot <- function(column, ylab) {
   return(qplot(data = df, x = 'simple',
                y = column, geom = 'boxplot',
-               xlab = ''))
+               xlab = '',
+               ylab = ylab))
 }
 
-grid.arrange(get_simple_boxplot(df$fixed.acidity),
-             get_simple_boxplot(df$volatile.acidity),
-             get_simple_boxplot(df$citric.acid),
-             get_simple_boxplot(df$TAC.acidity),
-             get_simple_boxplot(df$residual.sugar),
-             get_simple_boxplot(df$chlorides),
-             get_simple_boxplot(df$free.sulfur.dioxide),
-             get_simple_boxplot(df$total.sulfur.dioxide),
-             get_simple_boxplot(df$density),
-             get_simple_boxplot(df$pH),
-             get_simple_boxplot(df$sulphates),
-             get_simple_boxplot(df$alcohol),
+grid.arrange(get_simple_boxplot(df$fixed.acidity, 'fixed acidity'),
+             get_simple_boxplot(df$volatile.acidity, 'volatile acidity'),
+             get_simple_boxplot(df$citric.acid, 'citric acid'),
+             get_simple_boxplot(df$TAC.acidity, 'TAC acidity'),
+             get_simple_boxplot(df$residual.sugar, 'residual sugar'),
+             get_simple_boxplot(df$chlorides, 'chlorides'),
+             get_simple_boxplot(df$free.sulfur.dioxide, 'free sulf. dioxide'),
+             get_simple_boxplot(df$total.sulfur.dioxide, 'total sulf. dioxide'),
+             get_simple_boxplot(df$density, 'density'),
+             get_simple_boxplot(df$pH, 'pH'),
+             get_simple_boxplot(df$sulphates, 'sulphates'),
+             get_simple_boxplot(df$alcohol, 'alcohol'),
              ncol = 4)
@@
 ```{r echo=FALSE, Bivariate_Plots}
-get_bivariate_boxplot <- function(x, y, xlab, ylab) {
-  return(qplot(data = df, x = x, y = y, geom = 'boxplot'))
+get_bivariate_boxplot <- function(x, y, ylab) {
+  return(qplot(data = df, x = x, y = y, geom = 'boxplot', ylab = ylab))
 }
 
-grid.arrange(get_bivariate_boxplot(df$quality, df$fixed.acidity),
-             get_bivariate_boxplot(df$quality, df$volatile.acidity),
-             get_bivariate_boxplot(df$quality, df$citric.acid),
-             get_bivariate_boxplot(df$quality, df$TAC.acidity),
-             get_bivariate_boxplot(df$quality, log10(df$residual.sugar)),
-             get_bivariate_boxplot(df$quality, log10(df$chlorides)),
-             get_bivariate_boxplot(df$quality, df$free.sulfur.dioxide),
-             get_bivariate_boxplot(df$quality, df$total.sulfur.dioxide),
-             get_bivariate_boxplot(df$quality, df$density),
-             get_bivariate_boxplot(df$quality, df$pH),
-             get_bivariate_boxplot(df$quality, log10(df$sulphates)),
-             get_bivariate_boxplot(df$quality, df$alcohol),
+grid.arrange(get_bivariate_boxplot(df$quality, df$fixed.acidity,
+                                   'fixed acidity'),
+             get_bivariate_boxplot(df$quality, df$volatile.acidity,
+                                   'volatile acidity'),
+             get_bivariate_boxplot(df$quality, df$citric.acid,
+                                   'citric acid'),
+             get_bivariate_boxplot(df$quality, df$TAC.acidity,
+                                   'TAC acidity'),
+             get_bivariate_boxplot(df$quality, log10(df$residual.sugar),
+                                   'residual sugar'),
+             get_bivariate_boxplot(df$quality, log10(df$chlorides),
+                                   'chlorides'),
+             get_bivariate_boxplot(df$quality, df$free.sulfur.dioxide,
+                                   'free sulf. dioxide'),
+             get_bivariate_boxplot(df$quality, df$total.sulfur.dioxide,
+                                   'total sulf. dioxide'),
+             get_bivariate_boxplot(df$quality, df$density,
+                                   'density'),
+             get_bivariate_boxplot(df$quality, df$pH,
+                                   'pH'),
+             get_bivariate_boxplot(df$quality, log10(df$sulphates),
+                                   'sulphates'),
+             get_bivariate_boxplot(df$quality, df$alcohol,
+                                   'alcohol'),
              ncol = 4)
 
-grid.arrange(get_bivariate_boxplot(df$rating, df$fixed.acidity),
-             get_bivariate_boxplot(df$rating, df$volatile.acidity),
-             get_bivariate_boxplot(df$rating, df$citric.acid),
-             get_bivariate_boxplot(df$rating, df$TAC.acidity),
-             get_bivariate_boxplot(df$rating, log10(df$residual.sugar)),
-             get_bivariate_boxplot(df$rating, log10(df$chlorides)),
-             get_bivariate_boxplot(df$rating, df$free.sulfur.dioxide),
-             get_bivariate_boxplot(df$rating, df$total.sulfur.dioxide),
-             get_bivariate_boxplot(df$rating, df$density),
-             get_bivariate_boxplot(df$rating, df$pH),
-             get_bivariate_boxplot(df$rating, log10(df$sulphates)),
-             get_bivariate_boxplot(df$rating, df$alcohol),
+grid.arrange(get_bivariate_boxplot(df$rating, df$fixed.acidity,
+                                   'fixed acidity'),
+             get_bivariate_boxplot(df$rating, df$volatile.acidity,
+                                   'volatile acidity'),
+             get_bivariate_boxplot(df$rating, df$citric.acid,
+                                   'citric acid'),
+             get_bivariate_boxplot(df$rating, df$TAC.acidity,
+                                   'TAC acidity'),
+             get_bivariate_boxplot(df$rating, log10(df$residual.sugar),
+                                   'residual sugar'),
+             get_bivariate_boxplot(df$rating, log10(df$chlorides),
+                                   'chlorides'),
+             get_bivariate_boxplot(df$rating, df$free.sulfur.dioxide,
+                                   'free sulf. dioxide'),
+             get_bivariate_boxplot(df$rating, df$total.sulfur.dioxide,
+                                   'total sulf. dioxide'),
+             get_bivariate_boxplot(df$rating, df$density,
+                                   'density'),
+             get_bivariate_boxplot(df$rating, df$pH,
+                                   'pH'),
+             get_bivariate_boxplot(df$rating, log10(df$sulphates),
+                                   'sulphates'),
+             get_bivariate_boxplot(df$rating, df$alcohol,
+                                   'alcohol'),
              ncol = 4)
@@
 simple_cor_test <- function(x, y) {
-  return(cor.test(x, y)$estimate)
+  return(cor.test(x, as.numeric(y))$estimate)
 }
@@
 correlations <- c(
-  simple_cor_test(df$fixed.acidity, as.numeric(df$quality)),
-  simple_cor_test(df$volatile.acidity, as.numeric(df$quality)),
-  simple_cor_test(df$citric.acid, as.numeric(df$quality)),
-  simple_cor_test(df$TAC.acidity, as.numeric(df$quality)),
-  simple_cor_test(log10(df$residual.sugar), as.numeric(df$quality)),
-  simple_cor_test(log10(df$chlorides), as.numeric(df$quality)),
-  simple_cor_test(df$free.sulfur.dioxide, as.numeric(df$quality)),
-  simple_cor_test(df$total.sulfur.dioxide, as.numeric(df$quality)),
-  simple_cor_test(df$density, as.numeric(df$quality)),
-  simple_cor_test(df$pH, as.numeric(df$quality)),
-  simple_cor_test(log10(df$sulphates), as.numeric(df$quality)),
-  simple_cor_test(df$alcohol, as.numeric(df$quality)))
+  simple_cor_test(df$fixed.acidity, df$quality),
+  simple_cor_test(df$volatile.acidity, df$quality),
+  simple_cor_test(df$citric.acid, df$quality),
+  simple_cor_test(df$TAC.acidity, df$quality),
+  simple_cor_test(log10(df$residual.sugar), df$quality),
+  simple_cor_test(log10(df$chlorides), df$quality),
+  simple_cor_test(df$free.sulfur.dioxide, df$quality),
+  simple_cor_test(df$total.sulfur.dioxide, df$quality),
+  simple_cor_test(df$density, df$quality),
+  simple_cor_test(df$pH, df$quality),
+  simple_cor_test(log10(df$sulphates), df$quality),
+  simple_cor_test(df$alcohol, df$quality))
 names(correlations) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',

### 2015-01-28 p4/tester.py
add template files

@@ -1,0 +1,87 @@
+#!/usr/bin/pickle
+
+""" a basic script for importing student's POI identifier,
+    and checking the results that they get from it 
+ 
+    requires that the algorithm, dataset, and features list
+    be written to my_classifier.pkl, my_dataset.pkl, and
+    my_feature_list.pkl, respectively
+
+    that process should happen at the end of poi_id.py
+
+"""
+
+import pickle
+import sys
+sys.path.append("../tools/")
+from feature_format import featureFormat, targetFeatureSplit
+
+### load up student's classifier, dataset, and feature_list
+clf = pickle.load(open("my_classifier.pkl", "r") )
+dataset = pickle.load(open("my_dataset.pkl", "r") )
+feature_list = pickle.load(open("my_feature_list.pkl", "r"))
+
+### print basic info about the algorithm/parameters used
+print clf
+
+### prepare data for training/testing
+data = featureFormat(dataset, feature_list)
+labels, features = targetFeatureSplit(data)
+
+
+
+### stratified k-fold cross-validation is a form of 
+### CV where instances of each class are equally apportioned--
+### e.g. if you have 10% of one class and 90% of the other,
+### stratification means each fold will have 10% of one
+### class and 90% of the other
+###
+### this is helpful when you don't have a lot of instances
+### of one class or the other, because in that case the 
+### low-frequency class can become lopsided in the training-test
+### split skew the results
+from sklearn.metrics import precision_score
+from sklearn.metrics import recall_score
+from sklearn.cross_validation import StratifiedKFold
+skf = StratifiedKFold( labels, n_folds=3 )
+precisions = []
+recalls = []
+for train_idx, test_idx in skf: 
+    features_train = []
+    features_test  = []
+    labels_train   = []
+    labels_test    = []
+    for ii in train_idx:
+        features_train.append( features[ii] )
+        labels_train.append( labels[ii] )
+    for jj in test_idx:
+        features_test.append( features[jj] )
+        labels_test.append( labels[jj] )
+    
+    ### fit the classifier using training set, and test on test set
+    clf.fit(features_train, labels_train)
+    pred = clf.predict(features_test)
+
+
+    ### for each fold, print some metrics
+    print
+    print "precision score: ", precision_score( labels_test, pred )
+    print "recall score: ", recall_score( labels_test, pred )
+
+    precisions.append( precision_score(labels_test, pred) )
[…]

### 2015-01-29 p4/eda.py
add eda module

@@ -1,0 +1,49 @@
+#!/usr/bin/python
+import matplotlib.pyplot as plt
+import sys
+import pickle
+import csv
+sys.path.append("../tools/")
+
+from feature_format import featureFormat
+from feature_format import targetFeatureSplit
+
+def remove_keys(dict_object, keys):
+    """ removes a list of keys from a dict object """
+    for key in keys:
+        dict_object.pop(key, 0)
+
+def make_csv(data_dict):
+    """ generates a csv file from a data set """
+    fieldnames = ['name'] + data_dict.itervalues().next().keys()
+    with open('data.csv', 'w') as csvfile:
+        writer = csv.DictWriter(csvfile, fieldnames=fieldnames)
+        writer.writeheader()
+        for record in data_dict:
+            person = data_dict[record]
+            person['name'] = record
+            assert set(person.keys()) == set(fieldnames)
+            writer.writerow(person)
+
+def visualize(data_dict, feature_x, feature_y):
+    """ generates a plot of feature y vs feature x, colors poi """
+
+    data = featureFormat(data_dict, [feature_x, feature_y, 'poi'])
+
+    for point in data:
+        x = point[0]
+        y = point[1]
+        poi = point[2]
+        color = 'red' if poi else 'blue'
+        plt.scatter(x, y, color=color)
+    plt.xlabel(feature_x)
+    plt.ylabel(feature_y)
+    plt.show()
+
+if __name__ == '__main__':
+    data_dict = pickle.load(open("final_project_dataset.pkl", "r"))
+    outliers = ['TOTAL', 'THE TRAVEL AGENCY IN THE PARK', 'LOCKHART EUGENE E']
+    remove_keys(data_dict, outliers)
+    #make_csv(data_dict)
+    # visualize(data_dict, 'salary', 'bonus')
+    visualize(data_dict, 'from_poi_to_this_person', 'from_this_person_to_poi')

### 2015-01-29 p4/eda.py
restructure directories

@@ -1,49 +1,0 @@
-#!/usr/bin/python
-import matplotlib.pyplot as plt
-import sys
-import pickle
-import csv
-sys.path.append("../tools/")
-
-from feature_format import featureFormat
-from feature_format import targetFeatureSplit
-
-def remove_keys(dict_object, keys):
-    """ removes a list of keys from a dict object """
-    for key in keys:
-        dict_object.pop(key, 0)
-
-def make_csv(data_dict):
-    """ generates a csv file from a data set """
-    fieldnames = ['name'] + data_dict.itervalues().next().keys()
-    with open('data.csv', 'w') as csvfile:
-        writer = csv.DictWriter(csvfile, fieldnames=fieldnames)
-        writer.writeheader()
-        for record in data_dict:
-            person = data_dict[record]
-            person['name'] = record
-            assert set(person.keys()) == set(fieldnames)
-            writer.writerow(person)
-
-def visualize(data_dict, feature_x, feature_y):
-    """ generates a plot of feature y vs feature x, colors poi """
-
-    data = featureFormat(data_dict, [feature_x, feature_y, 'poi'])
-
-    for point in data:
-        x = point[0]
-        y = point[1]
-        poi = point[2]
-        color = 'red' if poi else 'blue'
-        plt.scatter(x, y, color=color)
-    plt.xlabel(feature_x)
-    plt.ylabel(feature_y)
-    plt.show()
-
-if __name__ == '__main__':
-    data_dict = pickle.load(open("final_project_dataset.pkl", "r"))
-    outliers = ['TOTAL', 'THE TRAVEL AGENCY IN THE PARK', 'LOCKHART EUGENE E']
-    remove_keys(data_dict, outliers)
-    #make_csv(data_dict)
-    # visualize(data_dict, 'salary', 'bonus')
-    visualize(data_dict, 'from_poi_to_this_person', 'from_this_person_to_poi')

### 2015-01-29 p4/poi_id.py

@@ -1,57 +1,0 @@
-#!/usr/bin/python
-
-import matplotlib.pyplot as plt
-import sys
-import pickle
-sys.path.append("../tools/")
-
-from feature_format import featureFormat
-from feature_format import targetFeatureSplit
-
-### features_list is a list of strings, each of which is a feature name
-### first feature must be "poi", as this will be singled out as the label
-features_list = ["poi"]
-
-
-### load the dictionary containing the dataset
-data_dict = pickle.load(open("final_project_dataset.pkl", "r") )
-
-### we suggest removing any outliers before proceeding further
-
-### if you are creating any new features, you might want to do that here
-### store to my_dataset for easy export below
-my_dataset = data_dict
-
-
-
-### these two lines extract the features specified in features_list
-### and extract them from data_dict, returning a numpy array
-data = featureFormat(my_dataset, features_list)
-
-
-
-### if you are creating new features, could also do that here
-
-
-
-### split into labels and features (this line assumes that the first
-### feature in the array is the label, which is why "poi" must always
-### be first in features_list
-labels, features = targetFeatureSplit(data)
-
-
-
-### machine learning goes here!
-### please name your classifier clf for easy export below
-
-clf = None    ### get rid of this line!  just here to keep code from crashing out-of-box
-
-
-### dump your classifier, dataset and features_list so 
-### anyone can run/check your results
-pickle.dump(clf, open("my_classifier.pkl", "w") )
-pickle.dump(data_dict, open("my_dataset.pkl", "w") )
-pickle.dump(features_list, open("my_feature_list.pkl", "w") )
-
-
-

### 2015-01-29 p4/scripts/eda.py

@@ -1,0 +1,49 @@
+#!/usr/bin/python
+import matplotlib.pyplot as plt
+import sys
+import pickle
+import csv
+sys.path.append("../tools/")
+
+from feature_format import featureFormat
+from feature_format import targetFeatureSplit
+
+def remove_keys(dict_object, keys):
+    """ removes a list of keys from a dict object """
+    for key in keys:
+        dict_object.pop(key, 0)
+
+def make_csv(data_dict):
+    """ generates a csv file from a data set """
+    fieldnames = ['name'] + data_dict.itervalues().next().keys()
+    with open('data.csv', 'w') as csvfile:
+        writer = csv.DictWriter(csvfile, fieldnames=fieldnames)
+        writer.writeheader()
+        for record in data_dict:
+            person = data_dict[record]
+            person['name'] = record
+            assert set(person.keys()) == set(fieldnames)
+            writer.writerow(person)
+
+def visualize(data_dict, feature_x, feature_y):
+    """ generates a plot of feature y vs feature x, colors poi """
+
+    data = featureFormat(data_dict, [feature_x, feature_y, 'poi'])
+
+    for point in data:
+        x = point[0]
+        y = point[1]
+        poi = point[2]
+        color = 'red' if poi else 'blue'
+        plt.scatter(x, y, color=color)
+    plt.xlabel(feature_x)
+    plt.ylabel(feature_y)
+    plt.show()
+
+if __name__ == '__main__':
+    data_dict = pickle.load(open("../data/final_project_dataset.pkl", "r"))
+    outliers = ['TOTAL', 'THE TRAVEL AGENCY IN THE PARK', 'LOCKHART EUGENE E']
+    remove_keys(data_dict, outliers)
+    #make_csv(data_dict)
+    # visualize(data_dict, 'salary', 'bonus')
+    visualize(data_dict, 'from_poi_to_this_person', 'from_this_person_to_poi')

### 2015-01-29 p4/scripts/poi_id.py

@@ -1,0 +1,57 @@
+#!/usr/bin/python
+
+import matplotlib.pyplot as plt
+import sys
+import pickle
+sys.path.append("../tools/")
+
+from feature_format import featureFormat
+from feature_format import targetFeatureSplit
+
+### features_list is a list of strings, each of which is a feature name
+### first feature must be "poi", as this will be singled out as the label
+features_list = ["poi"]
+
+
+### load the dictionary containing the dataset
+data_dict = pickle.load(open("../data/final_project_dataset.pkl", "r") )
+
+### we suggest removing any outliers before proceeding further
+
+### if you are creating any new features, you might want to do that here
+### store to my_dataset for easy export below
+my_dataset = data_dict
+
+
+
+### these two lines extract the features specified in features_list
+### and extract them from data_dict, returning a numpy array
+data = featureFormat(my_dataset, features_list)
+
+
+
+### if you are creating new features, could also do that here
+
+
+
+### split into labels and features (this line assumes that the first
+### feature in the array is the label, which is why "poi" must always
+### be first in features_list
+labels, features = targetFeatureSplit(data)
+
+
+
+### machine learning goes here!
+### please name your classifier clf for easy export below
+
+clf = None    ### get rid of this line!  just here to keep code from crashing out-of-box
+
+
+### dump your classifier, dataset and features_list so
+### anyone can run/check your results
+pickle.dump(clf, open("../data/my_classifier.pkl", "w") )
+pickle.dump(data_dict, open("../data/my_dataset.pkl", "w") )
+pickle.dump(features_list, open("../data/my_feature_list.pkl", "w") )
+
+
+

### 2015-01-29 p4/scripts/tester.py

@@ -1,0 +1,87 @@
+#!/usr/bin/pickle
+
+""" a basic script for importing student's POI identifier,
+    and checking the results that they get from it
+
+    requires that the algorithm, dataset, and features list
+    be written to my_classifier.pkl, my_dataset.pkl, and
+    my_feature_list.pkl, respectively
+
+    that process should happen at the end of poi_id.py
+
+"""
+
+import pickle
+import sys
+sys.path.append("../tools/")
+from feature_format import featureFormat, targetFeatureSplit
+
+### load up student's classifier, dataset, and feature_list
+clf = pickle.load(open("../data/my_classifier.pkl", "r") )
+dataset = pickle.load(open("../data/my_dataset.pkl", "r") )
+feature_list = pickle.load(open("../data/my_feature_list.pkl", "r"))
+
+### print basic info about the algorithm/parameters used
+print clf
+
+### prepare data for training/testing
+data = featureFormat(dataset, feature_list)
+labels, features = targetFeatureSplit(data)
+
+
+
+### stratified k-fold cross-validation is a form of
+### CV where instances of each class are equally apportioned--
+### e.g. if you have 10% of one class and 90% of the other,
+### stratification means each fold will have 10% of one
+### class and 90% of the other
+###
+### this is helpful when you don't have a lot of instances
+### of one class or the other, because in that case the
+### low-frequency class can become lopsided in the training-test
+### split skew the results
+from sklearn.metrics import precision_score
+from sklearn.metrics import recall_score
+from sklearn.cross_validation import StratifiedKFold
+skf = StratifiedKFold( labels, n_folds=3 )
+precisions = []
+recalls = []
+for train_idx, test_idx in skf:
+    features_train = []
+    features_test  = []
+    labels_train   = []
+    labels_test    = []
+    for ii in train_idx:
+        features_train.append( features[ii] )
+        labels_train.append( labels[ii] )
+    for jj in test_idx:
+        features_test.append( features[jj] )
+        labels_test.append( labels[jj] )
+
+    ### fit the classifier using training set, and test on test set
+    clf.fit(features_train, labels_train)
+    pred = clf.predict(features_test)
+
+
+    ### for each fold, print some metrics
+    print
+    print "precision score: ", precision_score( labels_test, pred )
+    print "recall score: ", recall_score( labels_test, pred )
+
+    precisions.append( precision_score(labels_test, pred) )
[…]

### 2015-01-29 p4/tester.py

@@ -1,87 +1,0 @@
-#!/usr/bin/pickle
-
-""" a basic script for importing student's POI identifier,
-    and checking the results that they get from it 
- 
-    requires that the algorithm, dataset, and features list
-    be written to my_classifier.pkl, my_dataset.pkl, and
-    my_feature_list.pkl, respectively
-
-    that process should happen at the end of poi_id.py
-
-"""
-
-import pickle
-import sys
-sys.path.append("../tools/")
-from feature_format import featureFormat, targetFeatureSplit
-
-### load up student's classifier, dataset, and feature_list
-clf = pickle.load(open("my_classifier.pkl", "r") )
-dataset = pickle.load(open("my_dataset.pkl", "r") )
-feature_list = pickle.load(open("my_feature_list.pkl", "r"))
-
-### print basic info about the algorithm/parameters used
-print clf
-
-### prepare data for training/testing
-data = featureFormat(dataset, feature_list)
-labels, features = targetFeatureSplit(data)
-
-
-
-### stratified k-fold cross-validation is a form of 
-### CV where instances of each class are equally apportioned--
-### e.g. if you have 10% of one class and 90% of the other,
-### stratification means each fold will have 10% of one
-### class and 90% of the other
-###
-### this is helpful when you don't have a lot of instances
-### of one class or the other, because in that case the 
-### low-frequency class can become lopsided in the training-test
-### split skew the results
-from sklearn.metrics import precision_score
-from sklearn.metrics import recall_score
-from sklearn.cross_validation import StratifiedKFold
-skf = StratifiedKFold( labels, n_folds=3 )
-precisions = []
-recalls = []
-for train_idx, test_idx in skf: 
-    features_train = []
-    features_test  = []
-    labels_train   = []
-    labels_test    = []
-    for ii in train_idx:
-        features_train.append( features[ii] )
-        labels_train.append( labels[ii] )
-    for jj in test_idx:
-        features_test.append( features[jj] )
-        labels_test.append( labels[jj] )
-    
-    ### fit the classifier using training set, and test on test set
-    clf.fit(features_train, labels_train)
-    pred = clf.predict(features_test)
-
-
-    ### for each fold, print some metrics
-    print
-    print "precision score: ", precision_score( labels_test, pred )
-    print "recall score: ", recall_score( labels_test, pred )
-
-    precisions.append( precision_score(labels_test, pred) )
[…]

### 2015-01-29 p4/tools/email_preprocess.py

@@ -1,0 +1,59 @@
+#!/usr/bin/python
+
+import pickle
+import numpy
+
+from sklearn import cross_validation
+from sklearn.feature_extraction.text import TfidfVectorizer
+from sklearn.feature_selection import SelectPercentile, f_classif
+
+
+
+def preprocess(words_file = "../tools/word_data.pkl", authors_file="../tools/email_authors.pkl"):
+    """
+        this function takes a pre-made list of email texts (by default word_data.pkl)
+        and the corresponding authors (by default email_authors.pkl) and performs
+        a number of preprocessing steps:
+            -- splits into training/testing sets (10% testing)
+            -- vectorizes into tfidf matrix
+            -- selects/keeps most helpful features
+
+        after this, the feaures and labels are put into numpy arrays, which play nice with sklearn functions
+
+        4 objects are returned:
+            -- training/testing features
+            -- training/testing labels
+
+    """
+
+    ### the words (features) and authors (labels), already largely preprocessed
+    ### this preprocessing will be repeated in the text learning mini-project
+    word_data = pickle.load( open(words_file, "r"))
+    authors = pickle.load( open(authors_file, "r") )
+
+    ### test_size is the percentage of events assigned to the test set (remainder go into training)
+    features_train, features_test, labels_train, labels_test = cross_validation.train_test_split(word_data, authors, test_size=0.1, random_state=42)
+
+
+
+    ### text vectorization--go from strings to lists of numbers
+    vectorizer = TfidfVectorizer(sublinear_tf=True, max_df=0.5,
+                                 stop_words='english')
+    features_train_transformed = vectorizer.fit_transform(features_train)
+    features_test_transformed  = vectorizer.transform(features_test)
+
+
+
+    ### feature selection, because text is super high dimensional and
+    ### can be really computationally chewy as a result
+    selector = SelectPercentile(f_classif, percentile=10)
+    selector.fit(features_train_transformed, labels_train)
+    features_train_transformed = selector.transform(features_train_transformed).toarray()
+    features_test_transformed  = selector.transform(features_test_transformed).toarray()
+
+    ### info on the data
+    print "no. of Chris training emails:", sum(labels_train)
+    print "no. of Sara training emails:", len(labels_train)-sum(labels_train)
+
+
+    return features_train_transformed, features_test_transformed, labels_train, labels_test

### 2015-01-29 p4/tools/parse_out_email_text.py

@@ -1,0 +1,54 @@
+#!/usr/bin/python
+
+from nltk.stem.snowball import SnowballStemmer
+import string
+
+def parseOutText(f):
+    """ given an opened email file f, parse out all text below the
+        metadata block at the top
+        (in Part 2, you will also add stemming capabilities)
+        and return a string that contains all the words
+        in the email (space-separated)
+
+        example use case:
+        f = open("email_file_name.txt", "r")
+        text = parseOutText(f)
+
+        """
+
+
+    f.seek(0)  ### go back to beginning of file (annoying)
+    all_text = f.read()
+
+    ### split off metadata
+    content = all_text.split("X-FileName:")
+    words = ""
+    if len(content) > 1:
+        ### remove punctuation
+        text_string = content[1].translate(string.maketrans("", ""), string.punctuation)
+
+        ### project part 2: comment out the line below
+        # words = text_string
+
+        ### split the text string into individual words, stem each word,
+        ### and append the stemmed word to words (make sure there's a single
+        ### space between each stemmed word)
+        stemmer = SnowballStemmer('english')
+        words_arr = map(stemmer.stem, text_string.split())
+        words = ' '.join(words_arr)
+
+
+    return words
+
+
+
+def main():
+    ff = open("../text_learning/test_email.txt", "r")
+    text = parseOutText(ff)
+    print text
+
+
+
+if __name__ == '__main__':
+    main()
+

### 2015-01-29 p4/tools/startup.py

@@ -1,0 +1,41 @@
+#!/usr/bin/python
+
+print
+print "checking for nltk"
+try:
+    import nltk
+except ImportError:
+    print "you should install nltk before continuing"
+
+print "checking for numpy"
+try:
+    import numpy
+except ImportError:
+    print "you should install numpy before continuing"
+
+print "checking for sklearn"
+try:
+    import sklearn
+except:
+    print "you should install sklearn before continuing"
+
+print
+print "downloading the Enron dataset (this may take a while)"
+print "to check on progress, you can cd up one level, then execute <ls -lthr>"
+print "Enron dataset should be last item on the list, along with its current size"
+print "download will complete at about 423 MB"
+import urllib
+url = "https://www.cs.cmu.edu/~./enron/enron_mail_20110402.tgz"
+urllib.urlretrieve(url, filename="../enron_mail_20110402.tgz") 
+print "download complete!"
+
+
+print
+print "unzipping Enron dataset (this may take a while)"
+import tarfile
+import os
+os.chdir("..")
+tfile = tarfile.open("enron_mail_20110402.tgz", "r:gz")
+tfile.extractall(".")
+
+print "you're ready to go!"

### 2015-01-29 p4/scripts/poi_id.py
remove outliers

@@ -8,6 +8,8 @@ sys.path.append("../tools/")
 from feature_format import featureFormat
 from feature_format import targetFeatureSplit
 
+import eda
+
 ### features_list is a list of strings, each of which is a feature name
 ### first feature must be "poi", as this will be singled out as the label
 features_list = ["poi"]
@@ -16,11 +18,14 @@ features_list = ["poi"]
 ### load the dictionary containing the dataset
 data_dict = pickle.load(open("../data/final_project_dataset.pkl", "r") )
 
-### we suggest removing any outliers before proceeding further
+### remove outliers
+outlier_keys = ['TOTAL', 'THE TRAVEL AGENCY IN THE PARK', 'LOCKHART EUGENE E']
+eda.remove_keys(data_dict, outlier_keys)
 
 ### if you are creating any new features, you might want to do that here
 ### store to my_dataset for easy export below
 my_dataset = data_dict
+
 
 
 

### 2015-01-29 p4/scripts/eda.py
add k-best features

@@ -1,49 +1,0 @@
-#!/usr/bin/python
-import matplotlib.pyplot as plt
-import sys
-import pickle
-import csv
-sys.path.append("../tools/")
-
-from feature_format import featureFormat
-from feature_format import targetFeatureSplit
-
-def remove_keys(dict_object, keys):
-    """ removes a list of keys from a dict object """
-    for key in keys:
-        dict_object.pop(key, 0)
-
-def make_csv(data_dict):
-    """ generates a csv file from a data set """
-    fieldnames = ['name'] + data_dict.itervalues().next().keys()
-    with open('data.csv', 'w') as csvfile:
-        writer = csv.DictWriter(csvfile, fieldnames=fieldnames)
-        writer.writeheader()
-        for record in data_dict:
-            person = data_dict[record]
-            person['name'] = record
-            assert set(person.keys()) == set(fieldnames)
-            writer.writerow(person)
-
-def visualize(data_dict, feature_x, feature_y):
-    """ generates a plot of feature y vs feature x, colors poi """
-
-    data = featureFormat(data_dict, [feature_x, feature_y, 'poi'])
-
-    for point in data:
-        x = point[0]
-        y = point[1]
-        poi = point[2]
-        color = 'red' if poi else 'blue'
-        plt.scatter(x, y, color=color)
-    plt.xlabel(feature_x)
-    plt.ylabel(feature_y)
-    plt.show()
-
-if __name__ == '__main__':
-    data_dict = pickle.load(open("../data/final_project_dataset.pkl", "r"))
-    outliers = ['TOTAL', 'THE TRAVEL AGENCY IN THE PARK', 'LOCKHART EUGENE E']
-    remove_keys(data_dict, outliers)
-    #make_csv(data_dict)
-    # visualize(data_dict, 'salary', 'bonus')
-    visualize(data_dict, 'from_poi_to_this_person', 'from_this_person_to_poi')

### 2015-01-29 p4/scripts/enron.py

@@ -1,0 +1,88 @@
+#!/usr/bin/python
+import csv
+import matplotlib.pyplot as plt
+import pickle
+import sys
+sys.path.append("../tools/")
+
+from feature_format import featureFormat
+from feature_format import targetFeatureSplit
+
+from sklearn.feature_selection import SelectKBest
+
+def remove_keys(dict_object, keys):
+    """ removes a list of keys from a dict object """
+    for key in keys:
+        dict_object.pop(key, 0)
+
+def make_csv(data_dict):
+    """ generates a csv file from a data set """
+    fieldnames = ['name'] + data_dict.itervalues().next().keys()
+    with open('data.csv', 'w') as csvfile:
+        writer = csv.DictWriter(csvfile, fieldnames=fieldnames)
+        writer.writeheader()
+        for record in data_dict:
+            person = data_dict[record]
+            person['name'] = record
+            assert set(person.keys()) == set(fieldnames)
+            writer.writerow(person)
+
+def visualize(data_dict, feature_x, feature_y):
+    """ generates a plot of feature y vs feature x, colors poi """
+
+    data = featureFormat(data_dict, [feature_x, feature_y, 'poi'])
+
+    for point in data:
+        x = point[0]
+        y = point[1]
+        poi = point[2]
+        color = 'red' if poi else 'blue'
+        plt.scatter(x, y, color=color)
+    plt.xlabel(feature_x)
+    plt.ylabel(feature_y)
+    plt.show()
+
+def get_k_best(data_dict, features_list, k):
+    """ runs scikit-learn's SelectKBest feature selection
+        returns dict where keys=features, values=scores
+    """
+    data = featureFormat(data_dict, features_list)
+    labels, features = targetFeatureSplit(data)
+
+    k_best = SelectKBest(k=k)
+    k_best.fit(features, labels)
+    scores = k_best.scores_
+    unsorted_pairs = zip(features_list[1:], scores)
+    sorted_pairs = list(reversed(sorted(unsorted_pairs, key=lambda x: x[1])))
+    k_best_features = dict(sorted_pairs[:k])
+    print "{0} best features: {1}".format(k, k_best_features.keys())
+    return k_best_features
+
+if __name__ == '__main__':
+    data_dict = pickle.load(open("../data/final_project_dataset.pkl", "r"))
[…]

### 2015-01-29 p4/scripts/poi_id.py

@@ -8,55 +8,78 @@ sys.path.append("../tools/")
 from feature_format import featureFormat
 from feature_format import targetFeatureSplit
 
-import eda
+import enron
 
-### features_list is a list of strings, each of which is a feature name
-### first feature must be "poi", as this will be singled out as the label
-features_list = ["poi"]
+# features_list is a list of strings, each of which is a feature name
+# first feature must be "poi", as this will be singled out as the label
+target_label = 'poi'
+email_features_list = [
+    # 'email_address', # remit email address; informational label
+    'from_messages',
+    'from_poi_to_this_person',
+    'from_this_person_to_poi',
+    'shared_receipt_with_poi',
+    'to_messages',
+    ]
+financial_features_list = [
+    'bonus',
+    'deferral_payments',
+    'deferred_income',
+    'director_fees',
+    'exercised_stock_options',
+    'expenses',
+    'loan_advances',
+    'long_term_incentive',
+    'other',
+    'restricted_stock',
+    'restricted_stock_deferred',
+    'salary',
+    'total_payments',
+    'total_stock_value',
+]
+features_list = financial_features_list + email_features_list
+for item in [features_list, email_features_list, financial_features_list]:
+    item.insert(0, target_label)
 
-
-### load the dictionary containing the dataset
+# load the dictionary containing the dataset
 data_dict = pickle.load(open("../data/final_project_dataset.pkl", "r") )
 
-### remove outliers
+# remove outliers
 outlier_keys = ['TOTAL', 'THE TRAVEL AGENCY IN THE PARK', 'LOCKHART EUGENE E']
 eda.remove_keys(data_dict, outlier_keys)
 
-### if you are creating any new features, you might want to do that here
-### store to my_dataset for easy export below
+# if you are creating any new features, you might want to do that here
+# store to my_dataset for easy export below
 my_dataset = data_dict
 
 
 
 
-### these two lines extract the features specified in features_list
-### and extract them from data_dict, returning a numpy array
+# these two lines extract the features specified in features_list
+# and extract them from data_dict, returning a numpy array
 data = featureFormat(my_dataset, features_list)
 
 
 
-### if you are creating new features, could also do that here
+# if you are creating new features, could also do that here
 
 
 
-### split into labels and features (this line assumes that the first
-### feature in the array is the label, which is why "poi" must always
-### be first in features_list
[…]

### 2015-01-29 p4/scripts/enron.py
add/remove features

@@ -58,6 +58,41 @@ def get_k_best(data_dict, features_list, k):
     print "{0} best features: {1}".format(k, k_best_features.keys())
     return k_best_features
 
+def add_poi_interaction(data_dict, features_list):
+    """ mutates data dict to add proportion of email interaction with pois """
+    fields = ['to_messages', 'from_messages',
+              'from_poi_to_this_person', 'from_this_person_to_poi']
+    for record in data_dict:
+        person = data_dict[record]
+        is_valid = True
+        for field in fields:
+            if person[field] == 'NaN':
+                is_valid = False
+        if is_valid:
+            total_messages = person['to_messages'] +\
+                             person['from_messages']
+            poi_messages = person['from_poi_to_this_person'] +\
+                           person['from_this_person_to_poi']
+            person['poi_interaction'] = float(poi_messages) / total_messages
+        else:
+            person['poi_interaction'] = 'NaN'
+    features_list += ['poi_interaction']
+
+def add_financial_aggregate(data_dict, features_list):
+    """ mutates data dict to add aggregate values from stocks and salary """
+    fields = ['total_stock_value', 'exercised_stock_options', 'salary']
+    for record in data_dict:
+        person = data_dict[record]
+        is_valid = True
+        for field in fields:
+            if person[field] == 'NaN':
+                is_valid = False
+        if is_valid:
+            person['financial_aggregate'] = sum([person[field] for field in fields])
+        else:
+            person['financial_aggregate'] = 'NaN'
+    features_list += ['financial_aggregate']
+
 if __name__ == '__main__':
     data_dict = pickle.load(open("../data/final_project_dataset.pkl", "r"))
     outliers = ['TOTAL', 'THE TRAVEL AGENCY IN THE PARK', 'LOCKHART EUGENE E']
@@ -86,3 +121,11 @@ if __name__ == '__main__':
                      'from_this_person_to_poi',
                      'shared_receipt_with_poi',
                      'to_messages']
+
+    add_poi_interaction(data_dict, features_list)
+    # visualize(data_dict, 'total_stock_value', 'poi_interaction')
+
+    add_financial_aggregate(data_dict, features_list)
+    # visualize(data_dict, 'poi_interaction', 'financial_aggregate')
+
+    k_best = get_k_best(data_dict, features_list, 10)

### 2015-01-29 p4/scripts/poi_id.py

@@ -1,5 +1,6 @@
 #!/usr/bin/python
 
+from copy import copy
 import matplotlib.pyplot as plt
 import sys
 import pickle
@@ -37,40 +38,34 @@ financial_features_list = [
     'total_payments',
     'total_stock_value',
 ]
-features_list = financial_features_list + email_features_list
-for item in [features_list, email_features_list, financial_features_list]:
-    item.insert(0, target_label)
+features_list = [target_label] + financial_features_list + email_features_list
 
 # load the dictionary containing the dataset
 data_dict = pickle.load(open("../data/final_project_dataset.pkl", "r") )
 
 # remove outliers
 outlier_keys = ['TOTAL', 'THE TRAVEL AGENCY IN THE PARK', 'LOCKHART EUGENE E']
-eda.remove_keys(data_dict, outlier_keys)
+enron.remove_keys(data_dict, outlier_keys)
 
-# if you are creating any new features, you might want to do that here
-# store to my_dataset for easy export below
-my_dataset = data_dict
+# instantiate copies of dataset and features for grading purposes
+my_dataset = copy(data_dict)
+my_feature_list = copy(features_list)
 
+# add two new features
+enron.add_financial_aggregate(my_dataset, my_feature_list)
+enron.add_poi_interaction(my_dataset, my_feature_list)
 
+# get K-best features
+best_features = enron.get_k_best(my_dataset, my_feature_list, 10).keys()
+my_feature_list = [target_label] + best_features
 
-
-# these two lines extract the features specified in features_list
-# and extract them from data_dict, returning a numpy array
-data = featureFormat(my_dataset, features_list)
-
-
-
-# if you are creating new features, could also do that here
-
-
+# extract the features specified in features_list
+data = featureFormat(my_dataset, my_feature_list)
 
 # split into labels and features (this line assumes that the first
 # feature in the array is the label, which is why "poi" must always
-# be first in features_list
+# be first in the features list
 labels, features = targetFeatureSplit(data)
-
-
 
 # machine learning goes here!
 # please name your classifier clf for easy export below

### 2015-01-29 p4/scripts/poi_id.py
add feature scaling

@@ -67,7 +67,14 @@ data = featureFormat(my_dataset, my_feature_list)
 # be first in the features list
 labels, features = targetFeatureSplit(data)
 
-# machine learning goes here!
+# scale features via min-max
+from sklearn import preprocessing
+scaler = preprocessing.MinMaxScaler()
+features = scaler.fit_transform(features)
+
+
+### machine learning classifier
+
 # please name your classifier clf for easy export below
 
 clf = None # get rid of this line!  just here to keep code from crashing out-of-box
@@ -76,5 +83,5 @@ clf = None # get rid of this line!  just here to keep code from crashing out-of-
 # dump your classifier, dataset and features_list so
 # anyone can run/check your results
 pickle.dump(clf, open("../data/my_classifier.pkl", "w"))
-pickle.dump(data_dict, open("../data/my_dataset.pkl", "w"))
-pickle.dump(features_list, open("../data/my_feature_list.pkl", "w"))
+pickle.dump(my_dataset, open("../data/my_dataset.pkl", "w"))
+pickle.dump(my_feature_list, open("../data/my_feature_list.pkl", "w"))

### 2022-02-19 udacity-full-stack/p3/application.py
Collapse p3

@@ -1,0 +1,331 @@
+import os
+from flask import Flask, flash, g, jsonify, redirect, render_template,\
+                  request, session, url_for
+from flask.ext.github import GitHub
+from flask.ext.sqlalchemy import SQLAlchemy
+
+app = Flask(__name__)
+app.config.from_object(os.environ['APP_SETTINGS'])
+
+# set-up github oauth - https://github-flask.readthedocs.org/en/latest/
+if os.environ.has_key('GITHUB_CLIENT_ID') and\
+   os.environ.has_key('GITHUB_CLIENT_SECRET'):
+    # set app configuration variables to environment variables
+    app.config['GITHUB_CLIENT_ID'] = os.environ['GITHUB_CLIENT_ID']
+    app.config['GITHUB_CLIENT_SECRET'] = os.environ['GITHUB_CLIENT_SECRET']
+else:
+    # set to None if they're not available
+    app.config['GITHUB_CLIENT_ID'] = None
+    app.config['GITHUB_CLIENT_SECRET'] = None
+github = GitHub(app)
+
+from database_setup import Base, User, Provider, Course
+from datetime import datetime
+from sqlalchemy import create_engine
+from sqlalchemy.orm import sessionmaker, scoped_session
+
+if True: #os.environ['APP_SETTINGS'] == "config.DevelopmentConfig":
+    engine = create_engine(os.environ['DATABASE_URL'])
+    Base.metadata.bind = engine
+
+    DBSession = scoped_session(sessionmaker(bind=engine))
+    db_session = DBSession()
+
+else:
+    app.config['SQLALCHEMY_DATABASE_URI'] = os.environ['DATABASE_URL']
+    db_session = SQLAlchemy(app)
+
+repo_uri = 'https://github.com/allanbreyes/mooc-catalog'
+base_uri = '/catalog/'
+api_uri = base_uri + 'api/'
+
+# oauth
+@app.route('/login')
+def login():
+    """ login route/routine """
+    if app.config['GITHUB_CLIENT_ID'] and app.config['GITHUB_CLIENT_SECRET']:
+        return github.authorize(redirect_uri=url_for('authorized', _external=True))
+    else:
+        flash('Howdy, developer! If you downloaded this, you still have to obtain a GitHub API client ID and key to get it wired up and working.', 'warning')
+        return redirect(url_for('index'))
+
+@app.route('/logout')
+def logout():
+    """ logout route/routine """
+    session.pop('user_id', None)
+    session.pop('user_token', None)
[…]

### 2022-02-19 udacity-full-stack/p3/config.py

@@ -1,0 +1,15 @@
+import os
+
+class Config(object):
+    DEBUG = False
+    TESTING = False
+    CSRF_ENABLED = True
+    SECRET_KEY = 'developmentkeyforthewin'
+    SQLALCHEMY_DATABASE_URI = os.environ['DATABASE_URL']
+
+class ProductionConfig(Config):
+    DEBUG = False
+
+class DevelopmentConfig(Config):
+    DEVELOPMENT = True
+    DEBUG = True

### 2022-02-19 udacity-full-stack/p3/database_setup.py

@@ -1,0 +1,68 @@
+import os
+import sys
+from sqlalchemy import Column, ForeignKey, Integer, String, Date, Boolean
+from sqlalchemy.ext.declarative import declarative_base
+from sqlalchemy.orm import relationship
+from sqlalchemy import create_engine
+
+Base = declarative_base()
+
+class User(Base):
+    __tablename__ = 'users'
+
+    id = Column(Integer, primary_key=True)
+    access_token = Column(String(200))
+
+    def __init__(self, access_token):
+        self.access_token = access_token
+
+class Provider(Base):
+    __tablename__ = 'providers'
+
+    id = Column(Integer, primary_key=True)
+    name = Column(String(80), nullable=False)
+    homepage_url = Column(String(255))
+
+    @property
+    def serialize(self):
+        """ JSON serializer method """
+        return {
+            'id': self.id,
+            'name': self.name,
+            'homepage_url': self.homepage_url
+        }
+
+class Course(Base):
+    __tablename__ = 'course'
+
+    id = Column(Integer, primary_key=True)
+    name = Column(String(80), nullable=False)
+    course_url = Column(String(255))
+    thumbnail_url = Column(String(255))
+    course_number = Column(String(20))
+    description = Column(String(2000))
+    start_date = Column(Date)
+    featured = Column(Boolean, default=False)
+    provider_id = Column(Integer, ForeignKey('providers.id'))
+    provider = relationship(Provider)
+    adder_id = Column(Integer, ForeignKey('users.id'))
+    adder = relationship(User)
+
+    @property
+    def serialize(self):
+        """ JSON serializer method """
+        return {
+            'id': self.id,
+            'name': self.name,
+            'course_url': self.course_url,
+            'thumbnail_url': self.thumbnail_url,
+            'course_number': self.course_number,
+            'description': self.description,
+            'start_date': str(self.start_date.isoformat()),
+            'featured': self.featured,
+            'provider_id': self.provider_id
+        }
+
+if __name__ == '__main__':
+    engine = create_engine(os.environ['DATABASE_URL'])
+    Base.metadata.create_all(engine)

### 2022-02-19 udacity-full-stack/p3/setup.sh

@@ -1,0 +1,13 @@
+# development/localhost
+export APP_SETTINGS="config.DevelopmentConfig"
+export DATABASE_URL="sqlite:///catalog.db"
+export GITHUB_CLIENT_ID=63af9832f4ddc627f1bc
+export GITHUB_CLIENT_SECRET=030eaa097a85af3e691156476d52fe29d4fa0bfa
+
+# production/heroku
+heroku config:set APP_SETTINGS=config.ProductionConfig --remote pro
+heroku config:set GITHUB_CLIENT_ID=###
+heroku config:set GITHUB_CLIENT_SECRET=###
+
+# production/apache2
+export DATABASE_URL="postgresql://localhost/catalog"

### 2022-02-19 udacity-full-stack/p3/vagrant/export_keys.sh

@@ -1,7 +1,0 @@
-# NOTE: Saving API keys to public repositories is a bad, bad idea.  However,
-# for the purposes of facilitating grading, they have been provided in this
-# script for export as environmental variables. The registered GitHub appli-
-# cation references localhost URIs and functions only, so security impact is
-# minimal.
-export GITHUB_CLIENT_ID="63af9832f4ddc627f1bc"
-export GITHUB_CLIENT_SECRET="030eaa097a85af3e691156476d52fe29d4fa0bfa"

### 2022-02-19 udacity-full-stack/p3/vagrant/pg_config.sh

@@ -1,12 +1,0 @@
-
-apt-get -qqy update
-apt-get -qqy install postgresql python-psycopg2
-apt-get -qqy install python-flask python-sqlalchemy
-apt-get -qqy install python-pip
-pip install bleach
-pip install GitHub-Flask
-pip install ipython
-su postgres -c 'createuser -dRS vagrant'
-su vagrant -c 'createdb'
-# su vagrant -c 'createdb catalog'
-# su vagrant -c 'psql catalog -f /vagrant/catalog/catalog.sql'

### 2022-02-19 data-science/p3/submission.Rmd
Rename nanodegree folders

--- 2015-01-26 p3/submission.Rmd
@@
 
-### Effect of acids on wine quality
+### Plot 1: Effect of acids on wine quality
 ```{r echo=FALSE, Plot_One}
@@
                                    fill = quality)) + 
-               ggtitle('Fixed Acidity in Different Wine Quality') +
                ylab('Fixed Acidity (g/dm^3)') +
@@
                                    fill = quality)) +
-               ggtitle('Volatile Acidity in Different Wine Quality') +
                ylab('Volatile Acidity (g/dm^3)') +
@@
                                    fill = quality)) +
-               ggtitle('Citric Acidity in Different Wine Quality') +
                ylab('Citric Acid (g/dm^3)') +
@@
                                    fill = quality)) +
-               ggtitle('pH in Different Wine Quality') +
                ylab('pH') +
@@
 
-### Effect of Alcohol on Wine Quality
+### Plot 2: Effect of Alcohol on Wine Quality
 ```{r echo=FALSE, Plot_Two}
@@
 
-### Plot Three
+### Plot 3: What makes good wines, good, and bad wines, bad?
 ```{r echo=FALSE, Plot_Three}

### 2022-02-19 data-science/p5/data/data.Rmd
Rename nanodegree folders

---
title: "Exploratory Data Analysis and Cleaning for RITA Flight Data"
author: "Allan Reyes"
date: "01/30/2015"
output: html_document
---

### About
This data set contains information on United States airline flight delays and performance.  The dataset includes all domestic flights from all carriers to/from major airports from June, 2003 through November, 2014.

[Download the data set from RITA.](http://www.transtats.bts.gov/OT_Delay/ot_delaycause1.asp?display=download&pn=0&month=11&year=2014)

### Load and inspect data
```{r}
# setwd('~/Dropbox/moocs/udacity-data-science/p5/data/')
df <- read.csv('334221194_112014_3544_airline_delay_causes.csv')
str(df)
summary(df)
head(df)
```

### Clean data
```{r}
library(dplyr)
# clean up date
df$date <- as.Date(paste(df$year, df$X.month, 1, sep='-'), format="%Y-%m-%d")
summary(df$date)

nrow(table(df$carrier))

# make a new summary table
ef <- df %>%
  group_by(date, year, carrier_name) %>%
  summarize(arrivals = sum(arr_flights),
            delayed = sum(arr_del15),
            cancelled = sum(arr_cancelled),
            diverted = sum(arr_diverted)) %>%
  transform(on_time = 1 - delayed/arrivals)

# stash NA values
ef <- ef[complete.cases(ef),]
```

### Exploratory Plots
```{r}
library(ggplot2)

ggplot(data = ef,
       aes(x = date, y = on_time)) +
  geom_line(aes(color = carrier_name))
```

### Find Airlines to Subset Data
```{r}
# aggregate by carrier name
agg <- ef %>%
  group_by(carrier_name) %>%
  summarize(monthly_avg = mean(arrivals),
            arrivals = sum(arrivals))

# pull 75th percentile, by monthly average arrivals
selected_carriers <- subset(agg, monthly_avg >= quantile(monthly_avg, 0.81))$carrier_name
selected_carriers

```

### Reshape Data
```{r}
ff <- subset(ef, is.element(carrier_name, selected_carriers)) %>%
  group_by(year, carrier_name) %>%
  summarize(arrivals = sum(arrivals),
            delayed = sum(delayed),
            cancelled = sum(cancelled),
            diverted = sum(diverted)) %>%
  transform(on_time = 1 - delayed/arrivals)
ff <- ff[complete.cases(ff),]

```

### Replot
```{r}
summary(df$year)

library(gridExtra)
p1 <- ggplot(data = ff,
       aes(x = year, y = on_time)) +
  geom_line(aes(color = carrier_name)) +
  scale_x_continuous(limits=c(2003, 2014), breaks=c(2003:2014))

p2 <- ggplot(data = ff,
       aes(x = year, y = arrivals)) +
  geom_line(aes(color = carrier_name)) +
  scale_x_continuous(limits=c(2003, 2014), breaks=c(2003:2014))

grid.arrange(p1, p2, ncol=1)
```

### Export New CSV
```{r}
write.csv(ff, file="data.csv", row.names=FALSE)
```
